\documentclass[a4paper,12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{tabularx}
\setlength{\parindent}{0pt}
\begin{document}

{\centering
\large \bf
CSE 201 Course Syllabus \\
Fall 2016 \\[\baselineskip]
}

Instructor: Dr. Turner \hfill Email: dturner@csusb.edu \\
Office: JB 350 \hfill Phone: (909) 537-5428 \\

\subsubsection*{Course Pages}

\begin{itemize}
\item Assignments: URL
\item Grade Sheet: URL
\end{itemize}

\subsubsection*{Office Hours}

...

\subsubsection*{Objective}

This course serves as an introduction to computer science and C++.

\subsubsection*{Prerequisites}

Some computer programming experience recommended.

\subsubsection*{Course Outline}
\begin{enumerate}

\item Data types and Basic I/O (Weeks 1 -- 2) 

Introduce the char, int, and double data types. Explain the common operators supported by these data types. Explain static arrays and C strings. Introduce standard I/O using cout and cin.

\item Conditions and Loops (Weeks 3 -- 4) 

Talk about boolean logic. Introduce comparison operators and the if statement. Introduce while, do while, and for statements. 

\item Functions and Pointers (Weeks 5 -- 6)

Show how to create functions, cover some useful functions from cmath, cstdlib, and ctime. Possibly cover pointers.

\item Classes and the C++ Library (Weeks 7 -- 8)

Show how to create classes. Cover some classes from the C++ library including string and vector.

\item Review (Week 9) 

Comprehensive review of topics covered thus far before final exam.

\end{enumerate}

\subsubsection*{Grading Policy}

Assignments (40\%), Quizzes (20\%), Final (40\%). No assignments accepted late. \\

Final grade is based on this table: \\

\begin{tabular}{|l|l|}
\hline
Percent & Grade \\
\hline
95 -- 100 & A \\
\hline
90 -- 94 & A- \\
\hline
87 -- 89 & B+ \\
\hline
84 -- 86 & B- \\
\hline
77 -- 79 & C+ \\
\hline
74 -- 76 & C \\
\hline
70 -- 73 & C- \\
\hline
67 -- 69 & D+ \\
\hline
64 -- 66 & D- \\
\hline
0 -- 59 & F \\
\hline
\end{tabular}

\subsubsection*{Course Schedule}

\begin{tabular}{|l|l|}
\hline
Week & Event \\
\hline
1 & Assignment 1 Due \\
\hline
2 & Quiz 1 \\
\hline
3 & Assignment 2 Due \\
\hline
4 & Quiz 2 \\
\hline
\end{tabular}

\end{document}
